\prefacesection{Acknowledgements}
Between my Dad giving me story problems as a kid and my Mom paying me for every A on my report card, they taught me to keep my mind busy.  Their praise for my academic achievements led me to getting a PhD.  Thanks to my sister Alexis too, I didn't forget you.

Mr. Schnepp taught me AP calculus in high school, but more importantly, to reason from basic principles.  I approach every problem with that mentality now.

Mr. Foy taught me AP chemistry, and to find humor in everything.

Professor Grbic provided me the opportunity to do research as an undergraduate alongside PhD students.  Carl Pfeiffer and Amit Patel somehow had the patience to teach me the fundamentals of electromagnetics, meta-materials, and tensor fields.

An NDSEG fellowship allowed me flexibility in my first 3 years of graduate school, for which I am very grateful.

Stephen Boyd's lectures organized my understanding of not just convex optimization, but pretty much all things math.

The excitement of Andrew Ng's machine learning course helped convince me to pursue AI.

At the deep learning reading group, I met and learned from some really smart people -- Andrej Karpathy, Richard Socher, Andrew Maas, and others.

Creative geniuses who shape the future, like my adviser Sebastian Thrun, are the reason I came to Stanford.  I don't think I could have had the same PhD experience at any other school or in any other lab.

I want to thank all the people on the skin cancer team -- Rob Novoa, Justin Ko, Helen Blau, Sebastian, Andre, and others.  It was exactly the kind of interdisciplinary collaboration that Stanford loves to brag about.  Being featured on the cover of Nature was pretty neat too.

I want to thank my beautiful girlfriend Akhila for encouraging me to power through this last mile of the degree.

Thank you to all the amazing people who I have been fortunate enough to learn from and who have invested their time and patience in me.  Someday when I’m older and wiser, I’ll pay it forward.